---
title: "Final Project"
author: "Neil Farrugia 17336831"
date: "14/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#https://www.kaggle.com/natashasavc/ie-students-winemag-analysis
#another idea is to look at the years!
#ie. year and price/points
library(tidyverse)
library(ggplot2)
library(tokenizers)
df <- read.csv("winemag-data-130k-v2.csv")
nas <- apply(apply(df, 2, is.na),2,sum)

barplot(nas,main="Missing values investigation",
        ylab="num of NAs",las=2)


on <- dim(df)[1]
vn <- dim(df)[2]
sum <- c("n observation"=on, "n variables" = vn)
print(sum)


```

From this brief missing values exploration, we can see that there is only 1 column that contains missing values - points column. With just over 8000 missing values.

The total number of observation is `r on` and the total number of variables is `r vn`.

Due to the fact that there is a high number of observation in this data set, completely ommitting the rows with a missing price will still allow plenty of data points to draw conclusions from.

```{r}
df1 <- df[complete.cases(df),]
```

```{r}
#range of the prices
summary(df1$price)
#Interqurtile range

#percentage under $200/250

tp <- table(df1$price)
p99 <- ceiling(length(df1$price)*0.995)
rngp99 <-(p99-20):(p99+20)
#tp99 <- tp[1:p99]
tpRS <- cumsum(tp)
pless <- tp[which(tpRS%in%rngp99)]
plessLabel <- labels(pless[length(pless)])

print(paste0("$1 to $",plessLabel," covers 99.5% of the total dataset"))
```


As you can see from the calculations above there is a really big range of prices. Due to the fact that under 99.5% of the wines in this data set are below $225, anything above 225 will be removed. Removing these outliers will allow for a better a visualisations of the analysis and distributions.

```{r}
df1 <- df1[which(df1$price < as.numeric(plessLabel)),]

```


```{r}

#
#lvl <- sort(unique(df1$points))
#df1$points <- factor(df1$points, levels = lvl)
#tabPo <- table(df1$price)

tabdf <- data.frame(table(df1$points))
tabdf <- tabdf[order(tabdf$Var1),]
avg <- mean(df1$points)
#mode <- tabdf$Var1[which(tabdf$Freq==max(tabdf$Freq))]
median <- median(80:100)
colors <- c("mean" = "green", "median" = "orange")
#colors <- c("mean" = "blue", "mode" = "red", "median" = "orange")
ggplot(df1,aes(x=points))+
  geom_histogram(binwidth = 1,aes(y=..density..))+
  #geom_vline(aes(xintercept = mode,color="mode"))+
  geom_vline(aes(xintercept = avg,color="mean"))+
  geom_vline(aes(xintercept = median,color="median"))+
  labs(title="Distribution of points",
         color = "Legend") +
 stat_function(fun=dnorm, args = list(mean = mean(df1$points), sd= sd(df1$points)), col = 'blue')+
 scale_color_manual(values = colors)

#ggplot(tabdf,aes(x=Var1,y=Freq))+
 # geom_histogram(stat='identity')+
#  geom_vline(aes(xintercept = mode,color="mode"))+
#  geom_vline(aes(xintercept = avg,color="mean"))+
#  geom_vline(aes(xintercept = median,color="median"))+
#  labs(title="Distribution of points",
#         y = "Points",
#         color = "Legend") +
# #stat_function(fun=dnorm, args = list(mean = mean(df1$points), sd= sd(df1$points)), col = 'blue')+
# scale_color_manual(values = colors)
#

avgP <- mean(df1$price)
#modeP <- mode(df1$price)
medianP <- median(df1$price) 
sumP <- c("mean Price" = avgP, "median Price" = medianP)


#bin = 10
#df1$price<-cut(df1$price,seq(0,max(df1$price),bin))
#price <- table(df1$price)
#price <- data.frame(price[c(1:(1000/bin))])

#look at wine prices, for different countries
#only look at top 10 countries with most wines
#only max 200euro
ctab <- sort(table(df1$country), decreasing = T)
ctab10 <- labels(ctab[1:10])[[1]]
df10 <- df1[which(df1$country%in%ctab10),]
ggplot(df1,aes(x=price))+
  geom_histogram(binwidth = 5,aes(y=..density..))+
  #geom_vline(aes(xintercept = mode,color="mode"))+
  geom_vline(aes(xintercept = avgP,color="mean"))+
  geom_vline(aes(xintercept = medianP,color="median"))+
  labs(title="Distribution of prices",
         color = "Legend") +
 stat_function(fun=dnorm, args = list(mean = mean(df1$price), sd= sd(df1$price)), col = 'blue')+
 scale_color_manual(values = colors)



ggplot(df1,aes(x = factor(1),y=log(price)))+
  geom_boxplot(fill = "red")+
  labs(title="Distribution of the log of Prices",
         y = "log(Price)", x = NULL) +
  annotate("text",label = paste("mean=",round(avgP,2)), x=1.3, y=4)
    #scale_color_manual(values = colors)+
    #theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

    
```
It was interesting to note the distribution of prices and points.
As you can see the points follow a normal distribution. Where as the prices the distrubtion is very much squewed to the right.

```{r}
ggplot(df1,aes(x = factor(1),y=log(price)))+
  geom_boxplot(fill = "red")+
  labs(title="Distribution of the log of Prices",
         y = "log(Price)", x = NULL) +
  annotate("text",label = paste("mean=",round(avgP,2)), x=1.3, y=4)


```


```{r}
#maybe look at price in different categories?
ggplot(df1,aes(x=points,y=price))+
  
  geom_point(position="jitter", alpha=1/10) + 
  geom_smooth(method="lm", se=F)
  

```
```{r}
df1$pgrp<-cut(df1$points,c(80,87,93,100),labels = c('low','medium','high'))
pvc <- table(data.frame(df1$country,df1$pgrp))
pvc <- addmargins(pvc)
pvcH <- data.frame(pvc[,'high']/pvc[,'Sum'])
pvcH$country<-rownames(pvcH)
colnames(pvcH)[1] <- "highPerc"

library(maps)
library(gapminder)
library(countrycode)
#library(countrycode,lib.loc = "/home/neilfa/R/x86_64-pc-linux-gnu-library/4.1")
world_map <- map_data("world")
wine_map <- left_join(pvcH, mosaicData::Countries, 
                           by = c("country" = "gapminder"))
wine_map <- left_join(wine_map, world_map, by = c("maps" = "region"))

as.factor(wine_map$country) %>% levels()
ggplot(wine_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = highPerc), color = "white") +
  scale_fill_viridis_c(option = "C")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
